---
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
# Diets

## Installation
You can install the diets package like this 

```
install.packages("remotes")
remotes::install_github("rhijmans/diets/pkg")
```

## Example run


### Nutrient content of food groups 

You can speficy a continent to adjust the proportion of items in a food goup, e.g. different vegetables and fruits.

```{r}
library(diets)
content <- nutrientContent(continent="Africa", redpalmoil=0.5, orangesweetpot=0.2)
head(content)
```

The table is in "long format" (all values are in one column). For each food group there are many "tags" that indicate the nutrients.
```{r}
unique(content$tag)
```

The unit is (generally) permille (1/1000)
```{r}
table(content$unit)
```

You can fortify the food. We provide some example data (it is real data). Here I show it for Senegal

```{r}
fort <- readRDS(system.file("ex/fortification.rds", package="diets"))
fort_sen <- fort[fort$area == "Senegal", ]
fort_sen

fcontent <- fortify(content, fort_sen)

# see the change
content[content$group=="Wheat and products" & content$tag=="ZN", ]
fort_sen[fort_sen$group=="Wheat and products" & fort_sen$tag=="ZN", ]
fcontent[fcontent$group=="Wheat and products" & fcontent$tag=="ZN", ]
```

### nutrient intake

First get conumption data. Here a FBS for Rwanda

```{r}
consumption <- readRDS(system.file("ex/rwa_FBS.rds", package = "diets"))
head(consumption)
consumption <- consumption[consumption$Element == "Food supply (kcal/capita/day)", c("Item", "Value")]
colnames(consumption) <- c("group", "value")
# consumption by food group in (kcal/capita) / day
head(consumption) 
```

Now combine that with the nutrient content data to to get nutrient intake.

```{r}
intake  <- nutrientIntake(consumption, content)
```


### adjust intake

You can adjust the intake for Ca, Fe and Zn 

Add Ca from water 

```{r}
intake <- add_Ca(intake)
```

Adjust Ca based on animal protein consumption
```{r}
intake <- adjust_Ca(intake)
```

Adjust Zn based for phytates based on Millar formula 
```{r}
intake <- adjust_Zn(intake)
```

Adjust Fe based on the amount of meat consumed
```{r}
iron <- readRDS(system.file("ex/NHemeIron.rds", package="diets"))
intake <- adjust_Fe(intake, iron)
```


### nutrient requirements

Get standard requirements 

```{r}
reqrda <- readRDS(system.file("ex/reqs_rda.rds", package = "diets"))
head(reqrda)
```

Get population data (by age, sex)

```{r}
pop <- readRDS(system.file("ex/rwa_pop.rds", package = "diets"))
head(pop)
```

Compute requirements for this population

```{r}
popreq <- nutrient_requirements(pop, reqrda)
head(popreq)
```

### deficiencies

Deficiencies are estimated using a coefficient of variation (if not provided, a default of 0.3 is used)

```{r}
CV <- readRDS(system.file("ex/CV.rds", package = "diets"))
head(CV)
```


And here we go:

```{r}
d <- deficiencies(intake, popreq, CV)
head(d)
```

Plot

```{r, fig.height=10}
par(mai=c(1,2,1,1))
barplot(d$deficiency, horiz=TRUE, names.arg=d$tag, las=1, cex.names=.66, col=rainbow(30))
```

I have not checked if this is correct (and why not...).

